### 2022-08-30 Manuscript_3_draft1.Rmd
Committed new version of preprint.

---
title: "Does differentiation beget differentiation?"
subtitle: Examining the link between implementation of and support for differentiation
thanks: "This is a very early draft. All comments are welcome at martin.moland@arena.uio.no. This version was finalized at `r format(Sys.time(), '%B %d, %Y')`."
author:
- name: Martin Moland
  affiliation: ARENA, Centre for European Studies
# bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`" # This would be more general
bibliography: inst/bib_paper_3.bib
csl: inst/european-journal-of-political-research.csl
abstract: "Questions about the public legitimacy of the EU’s increasingly flexible integration are attracting increasing scholarly interest. However, we still know little about whether exposure to past differentiation shapes public preferences for a more differentiated EU, and whether this varies according to how people think about EU membership. Combining recent survey data with data showing country-level exposure to differentiated integration, I show that those who have been exposed to more of it are likelier to want a future EU that allows countries to opt out of undesirable policy integration. This effect is particularly strong among Eurosceptics. My results suggest that exposure to the differentiated integration often used to integrate highly salient policy areas can solidify popular support for a less integrated EU, complicating the EU’s move towards an ‘ever closer Union’."
keywords: "differentiated integration, public opinion, Euroscepticism, European integration, contestation of the EU"
geometry: margin=1in
mainfont: cochineal
#sansfont: Linux Biolinum O
fontsize: 12pt
endnote: no
# pandocparas: TRUE
sansitup: FALSE
params:
  anonymous: ""
  doublespacing: ""
  removetitleabstract: ""
anonymous: "`r params$anonymous`"
doublespacing: "`r params$doublespacing`"
removetitleabstract: "`r params$removetitleabstract`"
header-includes:
  - \usepackage{longtable}
  - \LTcapwidth=.95\textwidth
  - \linespread{1.05}
  - \usepackage{hyperref}
  - \usepackage{setspace}
  - \usepackage{setspace}\doublespacing
---

```{r setup, include=FALSE}
is_docx <- knitr::pandoc_to("docx")
is_latex <- knitr::pandoc_to("latex")

# I don't know how Texas Instruments smart this is, but p-sure default DPI is 96.
# That's not a problem for LaTeX, but it looks not-so-great for Word.
# For Word, let's up that to 600. This should allow for cross-referencing 
# in {bookdown} while allowing for conditional DPI

if (is_latex) {
  conditional_dpi <- 300
  return(conditional_dpi)
} else {
  conditional_dpi <- 600
  return(conditional_dpi)
}

knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE, warning=FALSE,
                      fig.path='doc/figs/',
                      cache.path = 'doc/_cache/',
                      fig.width = 8.5, dpi = conditional_dpi,
                      fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      })

library(dplyr)
library(magrittr)
# library(tidyverse)
library(stevemisc)
library(modelsummary)
library(huxtable)
library(kableExtra)

allrds <- list.files(pattern = ".rds", path = "data") %>%
  stringr::str_remove(., ".rds")

for(i in allrds) {
      filepath <- file.path(paste0("data/",i,".rds"))
      assign(i, readRDS(filepath))
    }

```

# Introduction
Despite being among the world’s most advanced political cooperations, the expansion of the EU’s policy remit after the ratification of the Maastricht Treaty has made its integration ever-more differentiated [@leuffen_differentiated_2013; @schimmelfennig_ever_2020]. Internal EU differentiation typically takes two forms: First, some member states have opted out of cooperation to protect their own sovereignty. This is called “constitutional differentiation”. Second, through a process known as “instrumental differentiation”, countries with a diminished capacity for integration have been allowed to integrate at different schedules. This can happen either voluntarily or as a result of EU institutions imposing it as a precondition for membership [@schimmelfennig_instrumental_2014]. The differentiation of current European integration makes ‘variable geography’ [@stubb_categorization_1996] a better description of today’s EU than ‘ever closer Union’.

While increasing attention is being paid to why people want a differentiated EU [@leuffen_public_2022; @telle_elite_2022], we know little about how exposure to it has impacted support for a more differentiated future EU. While one of the few existing contributions examining the effect of differentiation on public opinion shows that it may strengthen the perceptions of the EU’s input legitimacy [@schraff_does_2020], two knowledge gaps remain: We still know too little about how past exposure to differentiation shapes support for it. Second, our understanding of how the relationship between exposure to differentiation and support for it is conditioned by attitudes towards the EU is still incomplete. Whether there is such a conditioning effect, either in the broader EU population, or where differentiation has been implemented in either a dominating or voluntary fashion is also unknown.

By combining survey data on citizens’ attitudes towards constitutional differentiated integration from 2020-2021 with data detailing differentiations from the EU’s treaties, I find that support for constitutional differentiation increases with exposure to differentiations from EU law, and that the extent to which this is the case strongly depends on your past attitudes towards the EU. However, I find that both where differentiation has been voluntary chosen, as in Scandinavia, or externally imposed in a discriminatory fashion, as in Central and Eastern Europe. While the reason for the generally greater support for differentiated integration among Eurosceptics in the countries that have been exposed to it may be that Eurosceptics are more likely than others to believe that differentiated integration is a path towards a less integrated Europe, more research is needed to understand why there is no real effect of pre-existing attitudes towards the EU in two regions with very different experiences of differentiation.

My findings are important to the discussion about the EU’s future spawned by Brexit [@european_commission_white_2017]: Because EU integration has expanded into increasingly salient areas, differentiation may be called upon to dampen public contestation of it [@schimmelfennig_ever_2020]. By providing greater numbers of citizens with exposure to differentiation, the EU’s institutions may solidify opposition to further integration among the Eurosceptics that have on several instances been able to put the brakes on further integration [@hooghe_postfunctionalist_2009].

Furthermore, the article contributes to the broader emerging literature on public support for international organizations [@dellmuth_legitimacy_2020; @dellmuth_elite_2021; @edwards_public_2009] While the EU’s legal and functional reach is much greater than that of other international organizations, there has been a turn towards expanding both the political and legal powers of many of the organizations constituting the global governance system [@zurn_theory_2018]. Even if more research is needed to understand whether my results travel to contexts other than the EU, my article helps shine a light on how public support for deeper political integration is impacted by non-uniform implementations of this emerging legal order.

# Public support for European integration - what we know and what we don't
A large literature investigates individual preferences for European integration [@basile_sharing_2020; @bremer_junckers_2020; @de_vries_variable_2013; @eichenberg_post-maastricht_2007; @gabel_public_1998; @hobolt_public_2015; @hooghe_calculation_2005]. The literature identifies three mechanisms that help shape attitudes towards the EU. The first is utilitarian, with support for the EU contingent on whether individuals or social groups with which they identify stand to gain from integration [@foster_economic_2021; @gabel_public_1998]. The second is cue-based, and posits that citizens use heuristics such as elite discourses, media messages, and national benchmarks to form opinions about European integration  [@harteveld_europe_2013; @marquart_knowing_2019; @steenbergen_whos_2007; @vries_euroscepticism_2018]. The literature also shows that strong identification with the nation-state makes citizens more likely to oppose EU integration [@borzel_identity_2020; @hooghe_calculation_2005; @luedtke_european_2005].

An under-investigated issue is why some people support a more differentiated EU. This gap is particularly pronounced in light of the large that literature already investigates differentiated integration conceptually [@batora_towards_2020; @leuffen_differentiated_2013; @schimmelfennig_instrumental_2014], empirically [@ganzle_differentiated_2019; @malang_political_2020; @ott_between_2017; @schimmelfennig_eastern_2017; @winzen_government_2020] and normatively [@bellamy_republican_2019; @eriksen_contesting_2019; @lord_autonomy_2021; @nicolaidis_new_2004]. The few contributions investigating individual support for differentiated integration has focused on liberal economic values [@leuffen_public_2022], Euroscepticism [de_blok_blessing_2021], territorially exclusive self-identification [@moland_opting_2022] and party cues [@telle_elite_2022].

Thus, with one notable exception, national contextual factors are under-studied in this literature. However, these contexts are crucial for explaining attitudes towards both the EU and other international organizations [@dellmuth_elite_2021; @edwards_public_2009; harteveld_europe_2013]. One contextual source of support for differentiated integration may be the exposure that citizens of EU countries have had to it: A country’s number of differentiations from EU law may bring its level of integration in and out of alignment with what a citizen would prefer. If there is a misalignment between the actual and desired levels of integration, citizens may respond by adjusting their stated preference for future differentiated integration [@busemeyer_beyond_2021; @wlezien_public_1995]. While previous literature investigates the potential feedback loops between EU policy-making and public opinion [@bolstad_dynamics_2015; @de_bruycker_democratically_2020] this paper shifts the focus to both how support for EU differentiation is impacted by exposure to differentiated integration generally and whether this effect is conditioned by attitudes towards European integration.

I believe that greater exposure to differentiated integration will increase support for it as a model of future integration. This is because such exposure is likely to familiarize citizens with EU differentiation, thus increasing their knowledge of how differentiated integration would work in the future. Greater knowledge is often, though not always, associated with greater support for European integration [@marquart_knowing_2019; @verhaegen_does_2015]. I furthermore find it theoretically plausible that whether this is the case will depend on whether one identifies with Eurosceptic sentiments or not: Because constitutional differentiated integration allows for permanent opt-outs from specific policy areas, Eurosceptics are more likely to see differentiated integration as a viable path towards the less differentiated Europe that they want to see in the future. This yields two complementary hypotheses:

*H1a: Exposure to differentiated integration increases support for future differentiated integration.*

```{r h1a, eval=!is_latex, results="asis"}
cat('> **H1a**: Exposure to differentiated integration increases support for future differentiated integration. ')
```

*H1b: The effect of exposure is larger among Eurosceptics than supporters of EU membership.*

```{r h1b, eval=!is_latex, results="asis"}
cat('> **H1b**: The effect of exposure is larger among Eurosceptics than supporters of EU membership.')
```

Responses to differentiated integration may further vary based on whether differentiation was imposed by the EU’s institutions or voluntarily chosen by member state government. An important reason for this is that citizens may perceive such differentiation as the EU dominating their countries, by shutting them out of economically beneficial integration, for instance in the field of freedom of movement [@cianciara_differentiated_2014; @schimmelfennig_instrumental_2014]. As @schimmelfennig_eu_2014 shows, the countries of Central and Eastern Europe were among those exposed to the greatest amount of discriminatory, rather than preferential, differentiated integration. The reason was that existing member states wanted to shield themselves from the economic destabilization believed to follow from the EU accession of countries like Poland, Estonia, and Hungary. As figure A6 shows, citizens of Central and Eastern Europe are among those in this sample exposed to the greatest amount of differentiated integration.

It is furthermore likely that Eurosceptics in these countries will have a more adverse reaction to differentiated integration than those who support EU membership: As differentiated integration in these countries has been predominantly discriminatory, they may be more likely to identify a non-uniformly integrated EU, in which countries can choose to opt out of integration if they so choose, with the potential for domination. Given the importance this group places on autonomy they are more likely to have adverse reactions to the discriminatory imposition of differentiated integration.

Elites in Central and Eastern Europe, especially in Poland, have also framed differentiated integration as a path towards a divided EU, with a powerful core and a less powerful periphery [@cianciara_differentiated_2014 178-179]. For example, Polish politicians opposed further integration of the Eurozone, a prominent example of differentiated integration, because they feared that the Central and Eastern member states, as non-members, would lose their influence over economic policy-making by not being inside the EU’s economic policy-making structures. Because elite cues are particularly important for shaping attitudes towards less salient issues, such as differentiated integration [@steenbergen_whos_2007; @telle_elite_2022], a reasonable assumption is that citizens would respond to these discourses by adjusting their preferences for integration downwards. Thus, two different causal mechanisms could produce the same outcome.

However, this history might be less relevant if it not being used by current actors. Instead, the increasingly Eurosceptic elite attitudes in Central and Eastern Europe [@borzel_identity_2020; @brack_towards_2020; @kriesi_politicization_2016] might form a more potent backdrop against which Central and Eastern European citizens might be expected to make up their minds. Exposure to such cues might, in contrast to what was stated before, make citizens more likely to express an increased demand for differentiated integration. This yields two competing hypotheses:

*H2a: Exposure to differentiated integration will lead to lower levels of support for future differentiated integration in Central and Eastern Europe than in the rest of the EU.*

```{r h2a, eval=!is_latex, results="asis"}
cat('> **H2a**: Exposure to differentiated integration in Central and Eastern will lead to lower levels of support for future differentiated integration than in the rest of the EU.')
```

*H2b: Exposure to differentiated integration in Central and Eastern Europe will lead to lower levels of support for it among Eurosceptics in these countries compared to supporters of EU membership.*

```{r h2b, eval=!is_latex, results="asis"}
cat('> **H2b**: Exposure to differentiated integration in Central and Eastern Europe will lead to lower levels of support for it among Eurosceptics in these countries compared to supporters of EU membership.')
```

*H3: Exposure to differentiated integration in Central and Eastern Europe will lead to higher levels of support for it than in the rest of the EU.*

```{r h3, eval=!is_latex, results="asis"}
cat('> **H3**: Exposure to differentiated integration in Central and Eastern Europe will lead to higher levels of support for it than in the rest of the EU.')
```

If exposure to discriminatorily imposed differentiation is assumed to lead to less support for differentiated integration, especially among Eurosceptics, the opposite is likely to be the case for countries with a great deal of voluntarily chosen differentiation. Scandinavia offers an interesting case for testing this assumption. Here the mode of implementation has been drastically different from that found in Central and Eastern Europe, as differentiation was both desired by national governments and justified through the autonomy-enhancing aspects of differentiated integration [@adler-nissen_opting_2014; @leruth_operationalizing_2015; @spendzharova_banking_2016]. Because of the previously explained cueing mechanism one might expect this pro-autonomy discourse to produce greater support for differentiated integration in Scandinavia than elsewhere in the EU.

Previous literature has found that the renewal of the Danish opt-out from justice and home affairs in 2015 led to stronger trust in EU democracy among Danish Eurosceptics [@schraff_does_2020]. This suggests both that citizens may respond positively to differentiation and that past exposure could make them more favourably inclined to it in the future. This is particularly likely to be true for Scandinavian Eurosceptics: These citizens may, because their countries are both relatively small and highly dependent on European market access, be more likely to see differentiation as a viable second-best alternative to the European disintegration they would otherwise favour. More generally it may also be the case that, because they are sceptical of EU integration, they are more likely to see differentiated integration as a goal in itself as it represents the kind of less integrated Europe that they would be likely to favour. I hypothesize the following:

*H4a: The positive effect of exposure to differentiated integration will be greater for citizens of Sweden and Denmark than in the EU as a whole.*

```{r h4a, eval=!is_latex, results="asis"}
cat('> **H4a**: The positive effect of exposure to differentiated integration will be greater for citizens of Sweden and Denmark than in the EU as a whole')
```

*H4b: The positive effect of exposure to differentiated integration will be greater for Scandinavian Eurosceptics than supporters of EU membership.*

```{r h4b, eval=!is_latex, results="asis"}
cat('> **H4b**: The positive effect of exposure to differentiated integration will be greater for Scandinavian Eurosceptics than supporters of EU membership.')
```

My paper thus investigates the impact of exposure to differentiation on two levels: I first investigate the general effect of exposure to differentiated integration in the EU as a whole, as well as in Scandinavia and Central and Eastern Europe. I then test whether this relationship is moderated by pre-existing Eurosceptic sentiments. Such a moderating effect is particularly likely to exist in either of the two regions because of the way that voluntary or discriminatory differentiation may be perceived as either an expansion or diminishing of the autonomy that this group is particularly likely to value.

# Data and methods
```{r load data}
EUI_final_DI_disc <- readRDS("data/EUI_final_DI_disc.Rds")
nordic_data <- readRDS("data/nordic_data.Rds")

```

I investigate these hypotheses using multilevel linear models and data from Yougov surveys fielded in 2020-21 [@hemerijck_eui-yougov_2021]. The sample includes respondents from Denmark, Finland, France, Germany, Sweden, Greece, Hungary, Italy, Lithuania, the Netherlands, Poland, Romania, and Spain and has an *N* of `r nrow(EUI_final_DI_disc)`. The surveys include two questions about differentiated integration. The first asks whether respondents want an EU where countries may choose to opt out of certain policy integration. This is called constitutional differentiation. The second asks whether the EU should allow countries to integrate at different speeds, while maintaining the goal of uniform integration in the long run. This is an instrumentally differentiated EU. This article focuses on popular support for the former, as its allowance for an EU that is permanently segmented into multiple policy communities represents arguably the sharpest departure from the EU’s normative goal of an “ever closer Union”.

I combine these data with a novel dataset [@schimmelfennig_eudiff1_2022] that features all differentiations from the EU treaties between 1952-2020. The data shows each instance of differentiation a country has been exposed to, the number of articles involved, and how long it was enforced. The dataset contains data on differentiations from EU treaties, or primary law, rather than differentiations from the less salient secondary law that constitutes the EU’s main legislative output [@duttle_opting_2017]. I use treaty differentiations because these are typically more publicly salient and subject to greater debate. Citizens are thus more likely to be aware of them than differentiations in secondary law.

## Dependent variable
My dependent variable measures support for constitutional differentiated integration. The variable uses a five-unit Likert scale and asks respondents whether they support a constitutionally differentiated EU (“Please tell us how far you agree or disagree with the following statement: Member states should be allowed to opt out of specific areas of European integration. This means that a member state can negotiate exceptions ("opt-out") for areas in which it does not wish to
cooperate. For example, Denmark has opted out of the common currency, and Poland has
opted out of the EU's Charter of Fundamental Rights.)". The response categories are ordered from 1-5, with 1 being Strongly agree and 5 being strongly disagree. I reverse the variable coding, so that higher values indicate agreement. By not asking about “constitutional differentiated integration”, the question becomes easier to understand. It also frames differentiation as a tool for strengthening autonomy and clarifies that countries can choose to opt out.

## Variables of key interest
To operationalize each country’s exposure to EU differentiation I multiply, for each differentiation, the number of articles that a country was exempt from with the number of years the differentiation was in force. This is then weighted based on the length of time that has passed since the differentiation ceased to be enforced, and finally log-transformed.

Operationalizing exposure to differentiation in this way has three main benefits: First, it captures both the temporal and functional scope of a country’s exposure to differentiation. Second, the weighting scheme means that those instances of differentiation that are still in force are given the most weight when investigating differentiation’s impact on support for differentiated integration in the future. This is likely to correspond to how respondents relate their own experience of differentiation to support for it: While many respondents from Central and Eastern Europe may still remember the particulars of their countries’ exposure to externally imposed differentiation, possibly producing a cohort effect among these individuals [@lauterbach_europe_2020], Southern European respondents are less likely to remember that they were exposed to externally imposed differentiation several decades ago and use this as an informational backdrop for their attitude-formation. Finally, the log-transformation reduces the influence of outliers.

To analyse the regionally specific effects of differentiation, I create dummies for Scandinavian and Central and Eastern European citizenship. I code all Swedish and Danish citizens with a 1 on the Scandinavian dummy. To create a Central and Eastern European dummy I do the same for everyone from Hungary, Lithuania, Poland, and Romania. These are used to investigate the interaction effects assumed by H2-H4. I also create a dummy for Euroscepticism, coding those who answer that they would vote yes to leaving the European Union in a hypothetical referendum with a 1 and those answering no with a 0.

I operationalize support for liberal economic policies, which has been found by previous literature to be a key determinant of support for differentiated integration [@leuffen_public_2022], through a question that asks whether respondents want to live in a Europe that “stresses economic integration, market competition and fiscal discipline”. This is contrasted to “a global Europe that acts as a leader on climate, human rights and global peace” or “a protective Europe that defends the European way of life and welfare against internal and external threats”. I code those who say that they prefer to live in a primarily market-oriented Europe as 1 and others 0 on this variable. While the question is multi-faceted, by asking about both fiscal discipline that may be described as austerity policies and more general market integration, I believe it still measures a general affinity for economic liberalism.

I include an equally weighted index of questions measuring respondents’ evaluation of the national economy. A factor analysis finds that these questions, in contrast to one asking about the respondent’s perceptions of the local job market, measure the same underlying dimension (see table A1). Previous literature shows that positive sociotropic economic evaluations correlate with support for both uniform and differentiated integration (see for instance @gabel_public_1998; @harteveld_europe_2013; @hooghe_calculation_2005; @leuffen_public_2022). Egotropic evaluations have, however, been shown to have little relevance for explaining support for differentiated integration [@leuffen_public_2022]. I measure support for national governments, relevant for explaining support for differentiated integration [@armingeon_loss_2014; @de_blok_blessing_2021; @harteveld_europe_2013], through a dummy asking respondents whether they believe their nation’s government is doing a good job.

I operationalize identity through a widely used question that asks respondents to rank the inclusiveness of their identities. I construct a dummy where those answering that they identify solely with their nation-states are coded as 1 and everyone else as 0. While nested identities combining a national and European component are common [@risse_euro_2003; @starke_concept_2021], previous literature finds the distinction between exclusively national and partially inclusive identities to be more important for producing Eurosceptic sentiments than variations between different nested identities [@hooghe_calculation_2005].

I also use ideology and gender as socio-political control variables. These are established predictors of support for both general and specific integration (see for example @carrubba_decision_2004; @hobolt_public_2015; @hooghe_calculation_2005; @schoen_identity_2008). I operationalise ideology through a 7-unit scale, with 0 representing far-left and 6 far-right positions. As previous literature also finds evidence of a curvilinear correlation between ideology and support for the EU [@toshkov_beyond_2022; @van_elsas_changing_2015], I add a squared indicator of left-right ideology.

```{r create descriptive data}
data_descriptives <- readRDS("data/Data_descriptives.Rds")

data_descriptives <- data_descriptives %>%
  rename(`Support constitutional DI` = opt_out_supp) %>%
  rename(`Article-years (logged)` = article_disc_log) %>%
  rename(`Liberal economic values` = market_europe) %>%
  rename(`Perception of economy` = alternative_econ) %>%
  rename(`Support national government` = gov_trust) %>%
  rename(`Left-right` = left_right) %>%
  rename(`Left-right (sqr.)` = left_right_sqr) %>%
  rename(`Eurosceptic` = eurosceptic) %>%
  rename(`Gender` = gender) %>%
  rename(`Exclusive identity` = excl_identity)
```

```{r}
library(modelsummary)

table_format <- ifelse(is_docx, "huxtable", 'default')

modelsummary::datasummary_skim(data_descriptives, fmt = 3, histogram = FALSE, 
                 title = "Descriptive statistics", output = "latex")

```

## Limitations
A limitation of the study is that both surveys were fielded during the COVID-19 pandemic, which could impact support for differentiation. However, @cicchi_eu_2020 show that solidarity, as is common, still seemed to reside at the national level in the early phase of the crisis. This means that my results may generalize to periods beyond those in which the survey was fielded. To account for how different baseline levels of support for differentiation are likely to exist in each country, I use multilevel models with random country intercepts.

There are also large amounts of missingness on certain variables (see table 1). Missing data may bias the results if they are correlated with individual-level characteristics like ideology or gender. I test the effect of such bias by multiply imputing the data [@buuren_mice_2011]. I show the results of this test in the section Robustness checks.

A further limitation, general to all studies using observational survey data, is the challenge of omitted variable bias. In this study omitted variable bias is particularly likely to stem from the absence of data on the respondents’ knowledge of the EU and education [@karp_know_2003; @marquart_knowing_2019]. I use sensitivity analysis [@cinelli_making_2020] to tentatively quantify the size of this bias and show the results of the analyses in the section Robustness checks.

# Results and discussion
I first show the effect of exposure to differentiation in the broader population, and especially how this interacts with already pre-existing Euroscepticism. I then discuss the effect of differentiation on support for it in both Scandinavia and Central and Eastern Europe.

```{r analysis 1}
library(lme4)
library(merTools)

```

```{r }

main_mod = list()

main_mod$"Euroscepticism" <- lmer(opt_out_supp ~ article_disc_log + market_europe + alternative_econ + gov_trust + left_right + left_right_sqr +
                                                         eurosceptic + gender + excl_identity + article_disc_log*eurosceptic + (1|country), weights = EUI_final_DI_disc$weight, 
                                                         data = EUI_final_DI_disc)

coef_map_eurosceptic <- c('(Intercept)' = 'Constant',
                          'article_disc_logeurosceptic1' = 'Article-years (logged) X Euroscepticism',
                          "article_disc_lognordic_dummy1" = "Article-years (logged) X Scandinavia", 
                          "article_disc_logCEE_dummy1" = "Article-years (logged) X CEE",
                          "nordic_dummy1" = "Scandinavia",
                          "CEE_dummy1" = "Central and Eastern Europe",
                          'article_disc_log' = 'Article-years (logged)',
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eurosceptic1' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'excl_identity' = 'Exclusively national identity')

modelsummary(main_mod, fmt = 3, coef_map = coef_map_eurosceptic, output = "latex", 
             stars = TRUE, title = "Model showing interaction between Euroscepticism and exposure to DI")
```

Both table 2 and figure 1 strongly support *H1a* and *H1b*, by confirming that exposure to differentiated integration is associated with greater support for differentiation as a future model for the EU. The effect of exposure is also significantly greater among Eurosceptics. The reason may be that Eurosceptic citizens in countries with greater exposure to differentiated integration are likelier to be familiar with the structure of a differentiated EU membership than those in countries not exposed to any differentiation. They may also see it as a more viable “second-best” alternative to full disintegration. However, even if the effect is larger for Eurosceptic than non-Eurosceptic citizens, the results suggest that future demands for differentiated integration among those most strongly exposed to it are unlikely to be restricted only to Eurosceptics.

```{r, fig.pos="H", fig.cap="Predicted probability of support for constitutional differentiation as a function of the interaction between article-years (logged) and Euroscepticism. 95% CIs."}

library(sjPlot)
library(ggplot2)
pred_plot_eurosceptic <- plot_model(main_mod$`Euroscepticism`, type = "pred", terms = c("article_disc_log", "eurosceptic"), 
                                    colors = "bw", legend.title = "Eurosceptic")

pred_plot_eurosceptic_final <-  pred_plot_eurosceptic + ggtitle("Support for constitutional differentiation", 
                                                                subtitle = "Support for differentiation among Eurosceptics and non-Eurosceptics") + labs(x = "Differentiation (article-years (logged))",
                                                                                                                     y = "Predicted value") + scale_linetype_manual(name = "Eurosceptic",
                           values = c("dashed", "solid", "dotted"),
                           labels = c("Non-Eurosceptic", "Eurosceptic"))

pred_plot_eurosceptic_final <- pred_plot_eurosceptic_final + font_size(title = 15, labels.x = 15, labels.y = 15)

pred_plot_eurosceptic_final
```

A plausible counter-argument to my hypothesized causal mechanism could be that differentiation was primarily chosen where underlying Euroscepticism, and thus a demand for differentiation, was already high. Figure A6 offers some support for such an assumption, as we find high levels of differentiated integration in Sweden and Denmark. These are countries where opt-outs from integration have been deployed as a response to a Euroscepticism shared by both political parties and voters [@leruth_operationalizing_2015; @spendzharova_banking_2016]. However, countries as diverse as Lithuania, Hungary, Italy, and the Netherlands have also seen differentiations from EU treaties. Differentiated integration is thus not a phenomenon restricted to the countries found in what @bolstad_dynamics_2015 identifies as a Eurosceptic periphery. This strengthens the case that the effect is not only driven by an endogenous relationship between Euroscepticism and differentiation.

I next analyse how support for constitutional differentiation is shaped by whether differentiation is voluntarily chosen or involuntarily imposed. I do this by comparing the interaction effects of Scandinavian and Central and Eastern European citizenship.

```{r }

main_regional = list()

main_regional$"Regional dummies" <- lmer(opt_out_supp ~ article_disc_log + market_europe + alternative_econ + gov_trust + left_right + left_right_sqr +
                                               eurosceptic + gender + excl_identity + nordic_dummy + CEE_dummy + article_disc_log*nordic_dummy + article_disc_log*CEE_dummy +  (1|country), 
                                               weights = EUI_final_DI_disc$weight, data = EUI_final_DI_disc)

library(modelsummary)
library(dplyr)
coef_map_regions <- c('(Intercept)' = 'Constant',
                          'article_disc_lognordic_dummy1' = 'Article-years (logged) X Scandinavia',
                      'article_disc_logCEE_dummy1' = 'Article-years (logged) X CEE',
                          'article_disc_log' = 'Article-years (logged)',
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eurosceptic1' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'excl_identity' = 'Exclusively national identity', 
                      "article_disc_logeurosceptic1" = "Article-years (logged) X Eurosceptic")

modelsummary(main_regional, fmt = 3, output = "latex", coef_map = coef_map_regions, 
             stars = TRUE, title = "Model showing regional interactions")

```

Table 3 shows a positive and statistically significant effect of the interaction between exposure to differentiated integration and Scandinavian citizenship. However, it finds no statistically significant effect of the interaction between Central and Eastern European citizenship and exposure to differentiated integration.

```{r, create plot}
library(sjPlot)
library(ggplot2)
#Scandinavia
pred_plot_nordics <- plot_model(main_regional$`Regional dummies`, type = "pred", terms = c("article_disc_log", "nordic_dummy"), 
                                legend.title = "Scandinavia", colors = "bw")

pred_plot_nordics_final <-  pred_plot_nordics + ggtitle("Scandinavia") + labs(x = "Differentiation (article-years (logged)",
                                                                                                                           y = "Predicted value")
pred_plot_nordics_final <- pred_plot_nordics_final + font_size(title = 15, labels.x = 15, labels.y = 15) + scale_linetype_manual(name = "Scandinavia",
                           values = c("dashed", "solid"),
                           labels = c("Non-Scandinavian", "Scandinavian"))

##CEE
pred_plot_CEE <- plot_model(main_regional$`Regional dummies`, type = "pred", terms = c("article_disc_log", "CEE_dummy"), 
                                    colors = "bw", legend.title = "CEE")

pred_plot_CEE_final <-  pred_plot_CEE + ggtitle("Central and Eastern Europe") + labs(x = "Differentiation (article-years (logged))",
                                                                                                                     y = "Predicted value") + scale_linetype_manual(name = "CEE",
                           values = c("dashed", "solid"),
                           labels = c("Non-CEE", "CEE"))

pred_plot_CEE_final <- pred_plot_CEE_final + font_size(title = 15, labels.x = 15, labels.y = 15)
```

```{r, fig.pos="H", fig.cap="Variations in support for differentiation as a function of exposure to DI (A = Scandinavia and B = CEE)", fig.height=8.5}
library(patchwork)

regional_final <- pred_plot_nordics_final/pred_plot_CEE_final + plot_annotation(tag_levels = "A", title = "Support for constitutional differentiation", subtitle = "Predicted values for regional interactions of exposure to differentiation in Scandinavia and Central and Eastern Europe", theme = theme(plot.title = element_text(size = 16)))

regional_final
```

Figure 2 supports this picture: Scandinavian citizens appear to become more favourably inclined towards differentiated integration as their exposure to it increases, compared to those from other regions. However, this effect is mainly present at the lower end of exposure to differentiation. Thus,I am able to confirm *H4a*, as distinct differences do exist within certain ranges of the variable. However, because of the large overlap for the largest range of the exposure variable this statistical significance must not be taken as indicative of a substantively large interaction effect.

Similarly, the predicted values show no significant difference between support for differentiation in Central and Eastern Europe and the rest of the EU. This leads me to reject *H2a*. I do, however, find some support for *H3*, whose underlying assumption is that exposure to Eurosceptic cues can lead to stronger support for differentiation even if it was imposed in a discriminatory fashion.

```{r}
region_mod <- readRDS("data/region_mod.Rds")

modelsummary::modelsummary(region_mod, fmt = 3, output = "latex", coef_map = coef_map_eurosceptic, 
             stars = TRUE, title = "Models showing interactions of Euroscepticism and region")
```

```{r create plot2}
library(sjPlot)
library(ggplot2)
#Scandinavia
pred_plot_nordics_eurosceptic <- plot_model(region_mod$Scandinavia, type = "pred", terms = c("article_disc_log", "eurosceptic"), 
                                legend.title = "Eurosceptic", colors = "bw")

pred_plot_nordics_eurosceptic_final <-  pred_plot_nordics_eurosceptic + ggtitle("Scandinavia") + labs(x = "Differentiation (article-years (logged))",
                                                                                                                           y = "Predicted value")
pred_plot_nordics_eurosceptic <- pred_plot_nordics_eurosceptic_final +
scale_linetype_manual(name = "Eurosceptic",
                           values = c("dashed", "solid"),
                           labels = c("Non-Eurosceptic", "Eurosceptic"))

##CEE
pred_plot_CEE_eurosceptic <- plot_model(region_mod$CEE, type = "pred", terms = c("article_disc_log", "eurosceptic"), 
                                legend.title = "Eurosceptic", colors = "bw")

pred_plot_CEE_eurosceptic_final <-  pred_plot_CEE_eurosceptic + ggtitle("Central and Eastern Europe") + labs(x = "Differentiation (article-years (logged))",
                                                                                                                           y = "Predicted value")
pred_plot_CEE_eurosceptic_final <- pred_plot_CEE_eurosceptic_final +  scale_linetype_manual(name = "Eurosceptic",
                           values = c("dashed", "solid"),
                           labels = c("Non-Eurosceptic", "Eurosceptic"))
```

```{r, fig.pos="H", fig.cap="Interaction between Euroscepticism and region. 95% CIs", fig.height=8.5}
library(patchwork)

regional_euroscepticism_final <- pred_plot_nordics_eurosceptic/pred_plot_CEE_eurosceptic_final + plot_annotation(tag_levels = "A", title = "Support for constitutional differentiation", subtitle = "Interaction of Euroscepticism and exposure to differentiation in CEE and Scandinavia", theme = theme(plot.title = element_text(size = 16)))

regional_euroscepticism_final
```

I must also reject *H2b*, as the effect of the interaction between differentiated integration and Euroscepticism is statistically insignificant in the Central and Eastern Europe data. I furthermore reject *H4b*, as there is no statistically significant difference between the effect of exposure for Eurosceptics and supporters of EU membership in Scandinavia. The lack of a lower level of support for differentiated integration among Central and Eastern Europeans, despite the fact that the discriminatory imposition of differentiation has been especially prevalent in these countries, suggests that historical trajectories and past experience with integration may predominantly lead to altered support for differentiation when it is used as a part of elite discourse. This finding is important for our understanding of how attitudes towards greater differentiation can be shaped by political elites in the future.

The fact that the surveys were fielded at the height of the COVID-19 pandemic means that we may be seeing lower levels of support for differentiated integration than we otherwise would: As @genschel_postfunctionalism_2021 argue, the COVID-19 pandemic “debordered” solidarity, making the prospect of mutual aid across member state borders more attractive to EU citizens. In times where EU solidarity is framed as a net positive for its citizens, support for differentiation may be lower than it otherwise would have been. More research is thus needed to establish whether the patterns established by this paper are consistent also in the future.

# Robustness checks
I use an approach outlined in @cinelli_making_2020 to test the sensitivity of several of my models to omitted variable bias (see figures A1-4). This quantifies the remaining variance that a potential confounder would need to account for to nullify the effect of theoretical interest, exposure to differentiated integration. A test of the sensitivity of the interaction between Euroscepticism and exposure to differentiation in Scandinavia (figure A3) shows that a variable explaining 50% of the remaining variance of support for constitutional differentiation would need to explain less than 10% of the remaining variance in real exposure to differentiation to nullify the effect of exposure. One potentially important omitted variable is knowledge of the EU [@leuffen_public_2022]. While knowledge of the EU is unlikely to be strongly connected with actual levels of differentiation, its strong connection to support for differentiation makes its omission particularly problematic. This suggests that omitted variable bias is likely to be an analytical concern throughout the entire study. Future studies should mitigate these concerns by including knowledge questions about current EU politics, to allow researchers to include similar controls as those commonly found in other EU public opinion research.

I also use multiple imputation through chained equations [@buuren_mice_2011; @fjelstul_explaining_2022] to test how sensitive the results are to bias caused by underlying patterns of data missingness. Multiple imputation uses each unit’s observed data to derive probable values for all covariates where a unit has missing data. This procedure is iterated many times, before the estimates derived from analysis of each imputed dataset are pooled. This procedure better accounts for the inherent uncertainty associated with data imputation and will thus yield more realistic parameter estimates and standard errors than traditional mean or zero imputation. As appendix tables A2-3 show, the multiply imputed models are similar to those found in tables 2-3. One important exception is the effect of exposure to differentiated integration in Central and Eastern Europe: While there is a weak positive effect of exposure in the main model, the multiply imputed model shows a weakly negative effect. However, both are statistically insignificant. Despite this minor deviation, the tests suggest that the effects established in this paper are not primarily driven by systematic data missingness.

I also test whether the substantial overlap between Scandinavian Eurosceptics and non-Eurosceptics is driven by the very low variability, with a standard deviation of `r round(sd(nordic_data$article_disc_log), digits = 2)`,  in the measure of exposure to differentiated integration in the Scandinavian data. I do this by testing a three-way interaction between Scandinavian citizenship, exposure to differentiated integration and Euroscepticism in the full data. Appendix table A6 suggests that the results are substantially similar to the interaction effect of Scandinavian citizenship and Euroscepticism shown in table. The results thus suggest that the estimates are not unduly biased by the very small standard deviation .

Lastly, I test how robust the results are to an alternative operationalization of Euroscepticism: satisfaction with EU democracy. Dissatisfaction with the state of EU democracy may be seen as a form of "regime scepticism", characterized by a dissatisfaction with EU decision-making, than "exit scepticism", as expressed through a desire to leave the union altogether [@vries_euroscepticism_2018]. Figure A5 ( coefficients in table A4) shows similar trends for both operationalizations: Those least satisfied with EU democracy, and who thus express the greatest amount of regime scepticism, also express the greatest support for constitutional differentiated integration. Their support for differentiation also increases more strongly as a function of exposure to previous differentiation than what is the case for those more satisfied with EU democracy. This strengthens the robustness of my original conclusions.

# Concluding discussion

This paper establishes, through a combination of data surveying citizen preferences for differentiated integration and their countries’ actual exposure to differentiation from the EU treaties, that there is a strong association between support for differentiated integration and past exposure to it. This may be because citizens more accustomed to differentiated integration potentially perceive it as a favourable way of combining the benefits of national membership of the EU with greater national autonomy. Importantly, I show that a potential tendency to see the kind of exposure that one has been exposed to as a preferable model for the future EU is, broadly speaking, conditioned by whether one opposes or supports EU membership. However, the results do not suggest that the mode of implementation of the differentiation that the same citizens have been exposed to play any large role in shaping attitudes towards a future differentiated EU.

What this suggests is that even if some differentiated integration could be perceived as a cartelization of the benefits of integration and as a dominating imposition on the political freedom of a polity, this does not necessarily translate to less support for differentiated integration as a policy. While my article does not establish why this is the case, it does strongly suggest that there is only a limited learning effect from exposure to differentiated integration. This is important because it implies that even the dominating imposition of a particular set of policies may not be enough to appreciably alter attitudes towards the same policy even among those who may be most likely to be critical of EU integration. Understanding what kind of discourses are more likely to produce critical attitudes towards differentiated integration in this group is an important question for future research.

Another question is whether the public is equally likely to prefer differentiation in all policy areas. Existing research [@de_blok_blessing_2021] has largely focused on individual support for differentiated integration of the EU as a polity. It is both possible and probable that even those favouring an “ever closer Union” may have diverging preferences for integration at the level of individual policies. Testing where individual citizens may prefer more differentiation may also help theorize the ambivalent attitudes towards EU integration found by an increasingly prevalent strand of the EU public opinion literature [@de_vries_variable_2013; @goldberg_one_2021; @moland_constraining_2022].

My results have policy implications for both the EU and other international organizations seeking to deepen their own political integration. They suggest first that allowing for a greater institutionalization of differentiated integration may undermine the public legitimacy of the EU’s goal of ‘ever closer Union’. However, beyond the EU, the results also suggest that the public legitimacy of deeper political integration may be challenged if citizens gradually become accustomed to more differentiated and non-uniform forms of integration. While one should be wary of generalizing findings from the EU to other international organizations, my results suggest that allowing some countries to opt out of the legal order that is increasingly developing beyond the EU [@zurn_theory_2018] may lead to greater popular opposition to the idea that countries should be members of entire frameworks of political cooperation, rather than more functionally limited partnerships for solving particular problems.

# References
<div id="refs"></div>

# Appendix
\setcounter{table}{0}
\renewcommand*{\thetable}{A\arabic{table}}
```{r, results='asis'}

appendix_items <- readRDS("data/appendix_items.Rds")
library(psychTools)

cat(fa2latex(appendix_items$`Factor analysis of economy`, heading = "Factor analysis of perception of economy", silent = TRUE, caption = "Factor analysis of perceptions of the economy."))
```
\setcounter{figure}{0}
\renewcommand*{\thefigure}{A\arabic{figure}}
```{r, fig.cap="Sensitivity to omitted variable bias for article-years (logged)", fig.height=8.5}
library(sensemakr)
appendix_items$"Fixed effect of non-interacted model" <- lm(opt_out_supp ~ article_disc_log + alternative_econ + gov_trust + left_right + left_right_sqr +
                          eurosceptic + gender + excl_identity + factor(country), weights = EUI_final_DI_disc$weight, 
                        data = EUI_final_DI_disc)

#Run sensemakr
appendix_items$"Sensitivity test of non-interacted model" <- sensemakr::sensemakr(appendix_items$`Fixed effect of non-interacted model`, 
                                treatment = "article_disc_log")

plot(appendix_items$"Sensitivity test of non-interacted model", type = "extreme")

```

```{r, fig.cap="Sensitivity to omitted variable bias for article-years (logged) X Euroscepticism", fig.height=8.5}

appendix_items$"Fixed effect of interacted model" <- lm(opt_out_supp ~ article_disc_log + alternative_econ + gov_trust + left_right + left_right_sqr +
                        eurosceptic + gender + excl_identity + article_disc_log*eurosceptic + factor(country), weights = EUI_final_DI_disc$weight, 
                      data = EUI_final_DI_disc)

appendix_items$"Sensitivity test of interacted model" <- sensemakr(appendix_items$`Fixed effect of interacted model`, 
                                  treatment = "article_disc_log:eurosceptic1")

plot(appendix_items$`Sensitivity test of interacted model`, type = "extreme")

```

```{r, fig.cap="Sensitivity to omitted variable bias for article-years (logged) X Euroscepticism in Scandinavia", fig.height=8.5}
nordic_data <- readRDS("data/nordic_data.Rds")
CEE_data <- readRDS("data/CEE_data.Rds")

appendix_items$"Fixed effect of interacted model (Nordics)" <- lm(opt_out_supp ~ article_disc_log + alternative_econ + gov_trust + left_right + left_right_sqr +
                                                          eurosceptic + gender + excl_identity + article_disc_log*eurosceptic + factor(country), weights = nordic_data$weight, 
                                                        data = nordic_data)

appendix_items$"Sensitivity test of interacted model (Nordics)" <- sensemakr(appendix_items$`Fixed effect of interacted model (Nordics)`, 
                                                                   treatment = "article_disc_log:eurosceptic1")

plot(appendix_items$`Sensitivity test of interacted model (Nordics)`, type = "extreme")
```

```{r, fig.cap="Sensitivity to omitted variable bias for article-years (logged) X Euroscepticism in Central and Eastern Europe", fig.height=8.5}
library(sensemakr)
appendix_items$"Fixed effect of interacted model (CEE)" <- lm(opt_out_supp ~ article_disc_log + alternative_econ + gov_trust + left_right + left_right_sqr +
                                                                    eurosceptic + gender + excl_identity + article_disc_log*eurosceptic + factor(country), weights = CEE_data$weight, 
                                                                  data = CEE_data)

appendix_items$"Sensitivity test of interacted model (CEE)" <- sensemakr(appendix_items$`Fixed effect of interacted model (CEE)`, 
                                                                             treatment = "article_disc_log:eurosceptic1")

plot(appendix_items$`Sensitivity test of interacted model (CEE)`, type = "extreme")

```

```{r}

library(mice)
library(miceadds)
library(merTools)
main_impute <- readRDS("data/main_impute.Rds")
```

```{r}
coef_map_regions2 <- c('(Intercept)' = 'Constant',
                          'article_disc_log:nordic_dummy' = 'Article-years (logged) X Scandinavia',
                      'article_disc_log:CEE_dummy' = 'Article-years (logged) X CEE',
                      "article_disc_log:eurosceptic1" = "Article-years (logged) X Eurosceptic",
                          'article_disc_log' = 'Article-years (logged)',
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eurosceptic1' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'excl_identity' = 'Exclusively national identity', 
                      "nordic_dummy" = "Scandinavia",
                      "CEE_dummy" = "Central and Eastern Europe")

modelsummary(main_impute, fmt = 3, output = "latex", coef_map = coef_map_regions2, 
             stars = TRUE, title = "Imputed model of interaction between Euroscepticism and article-years (logged) and model with regional dummies")
```

```{r}
regional_impute <- readRDS("data/regional_impute.Rds")

coef_map_eurosceptic2 <- c('(Intercept)' = 'Constant',
                          'article_disc_log:eurosceptic1' = 'Article-years (logged) X Euroscepticism',
                          'article_disc_log' = 'Article-years (logged)',
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eurosceptic1' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'excl_identity' = 'Exclusively national identity')

modelsummary::modelsummary(regional_impute, fmt = 3, output = "latex", coef_map = coef_map_eurosceptic2, 
             stars = TRUE, title = "Imputed models showing interactions of Euroscepticism and article-years (logged) by region")
```

```{r}

coef_map_alteurosceptic <- c('(Intercept)' = 'Constant',
                          'article_disc_logeu_democ_satisf' = 'Article-years (logged) X Euroscepticism',
                          'article_disc_log' = 'Article-years (logged)',
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eu_democ_satisf' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'excl_identity' = 'Exclusively national identity')

modelsummary::modelsummary(appendix_items$`Alternative conceptualization of Euroscepticism`, fmt = 3, output = "latex", coef_map = coef_map_alteurosceptic, 
             stars = TRUE, title = "Model showing alternative conceptualization of Euroscepticism")
```

```{r, fig.cap="Predicted values of support for constitutional differentiated integration as a function of an interaction between exposure to differentiated integration and satisfaction with EU democracy. Slopes for minimum, median and maximum values of satisfaction with EU democracy are shown. 95% CIs", fig.height=8.5}

pred_plot_interact_main <- plot_model(appendix_items$`Alternative conceptualization of Euroscepticism`, type = "pred", terms = c("article_disc_log", "eu_democ_satisf"), 
                                      colors = "bw", legend.title = "EU democracy satisfaction")

pred_plot_interact_main_final <-  pred_plot_interact_main + ggtitle("Support for constitutional differentiation", 
                                                                    subtitle = "Interaction between EU democracy satisfaction and differentiation") + labs(x = "Differentiation (article-years (logged))",
                                                                                                                                                           y = "Predicted value") + scale_linetype_manual(name = "EU democracy satisfaction",
                           values = c("dashed", "solid", "dotted"),
                           labels = c("Minimum", "Median", "Max")) + font_size(title = 15, labels.x = 15, labels.y = 15)

pred_plot_interact_main_final
```

```{r, fig.cap="Number of differentiations from EU primary law 1952-2020 (log-transformed and recency-weighted article-years) by EU member state.", fig.height=8.5}
bar_chart_articleyear <- ggplot(EUI_final_DI_disc, aes(x = factor(country), y = article_disc_log)) +
  geom_col(stat = "identity", position = "dodge")

bar_chart_articleyear + labs(x = "Country", y = "Discounted article-years (log-transformed)") +
  scale_x_discrete(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14),
                   labels = c("UK", "DK", "FI", "FR", "DE", "SE", "GR", "HU", "IT",
                              "LT", "NL", "PL", "RO", "ES")) +
  ggtitle(label = "Differentiated integration in the EU", 
          subtitle = "Primary law differentiations (article-years log-transformed and weighted)") + font_size(title = 15, labels.x = 15, labels.y = 15)
```

```{r}
southern_mod = list()

southern_mod$"Southern Europe" <- lmer(opt_out_supp ~ article_disc_log + market_europe + alternative_econ + gov_trust + left_right + left_right_sqr +
                           eurosceptic + gender + excl_identity + south_dummy + south_dummy*article_disc_log + (1|country), weights = EUI_final_DI_disc$weight, 
                         data = EUI_final_DI_disc)

coef_map_southern <- c('(Intercept)' = 'Constant',
                          'article_disc_logsouthern_dummy1' = 'Article-years (logged) X Southern Europe',
                          'article_disc_log' = 'Article-years (logged)',
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eurosceptic1' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'excl_identity' = 'Exclusively national identity', 
                      "south_dummy1" = "Southern Europe")

modelsummary::modelsummary(southern_mod, fmt = 3, output = "latex", coef_map = coef_map_alteurosceptic, 
             stars = TRUE, title = "Model showing effect of exposure in Southern Europe")

```

```{r, fig.pos="H", fig.cap= "Support for differentiated integration in Southern Europe as an interaction between Euroscepticism and exposure to differentiated integration. 95% CIs.", fig.height=8.5}

pred_plot_southern <- plot_model(southern_mod$`Southern Europe`, type = "pred", terms = c("article_disc_log", "south_dummy"), 
                                 colors = "bw", legend.title = "Southern Europe")

pred_plot_southern_final <-  pred_plot_southern + ggtitle("Support for constitutional differentiation (Southern Europe)", subtitle = "Support for differentiation among Southern and non-Southern Europeans") + labs(x = "Differentiation (article-years (logged))",
                                                                                                                                 y = "Predicted value")
pred_plot_southern_final + font_size(title = 15, labels.x = 15, labels.y = 15) + scale_linetype_manual(name = "Southern Europe",
                           values = c("dashed", "solid"),
                           labels = c("Non-Southern", "Southern"))
```

```{r }
scandi_eurosceptic = list()

scandi_eurosceptic$"Scandinavia (three-way interactions)" <- lmer(opt_out_supp ~ article_disc_log + market_europe + alternative_econ + gov_trust + left_right + left_right_sqr +
                             eurosceptic + gender + excl_identity + nordic_dummy + article_disc_log*eurosceptic*nordic_dummy + (1|country), weights = EUI_final_DI_disc$weight, 
                           data = EUI_final_DI_disc)

coef_map_nordic <- c('(Intercept)' = 'Constant',
                          'article_disc_logeurosceptic1nordic_dummy1' = 'Article-years (logged) X Eurosceptic X Scandinavia',
                          'article_disc_log' = 'Article-years (logged)',
                          'nordic_dummy1' = "Scandinavia",
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eurosceptic1' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'excl_identity' = 'Exclusively national identity')

modelsummary::modelsummary(scandi_eurosceptic, fmt = 3, output = "latex", coef_map = coef_map_nordic,
             stars = TRUE, title = "Model showing three-way interaction between Scandinavian dummy, Eurosceptic and article-years (logged)")
```