### 2022-11-02 appendix.Rmd

---
title: "Online supplementary material: 'Does differentiation beget differentiation?'"
author:
- name: Martin Moland
  affiliation: ARENA, Centre for European Studies
geometry: margin=1in
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
output: pdf_document
keywords: ""
endnotes: FALSE
# csl: inst/jpr.csl
mainfont: cochineal
# sansfont: Linux Biolinum O
sansitup: FALSE
params:
  anonymous: TRUE
  doublespacing: ""
  removetitleabstract: ""
appendix: TRUE
appendixletter: "A"
pandocparas: FALSE
anonymous: "`r params$anonymous`"
doublespacing: "`r params$doublespacing`"
removetitleabstract: "`r params$removetitleabstract`"
linkcolor: black
header-includes:
  - \usepackage{longtable}
  - \usepackage{csquotes}
  - \LTcapwidth=.95\textwidth
  - \linespread{1.05}
  - \usepackage{hyperref}
  - \usepackage{setspace}
  - \renewcommand{\thepage}{A-\arabic{page}}
  - \usepackage{booktabs}
  - \usepackage{siunitx}
  - \newcolumntype{d}{S[input-symbols = ()]}
---

```{r setup, include=FALSE}
is_docx <- knitr::pandoc_to("docx")
is_latex <- knitr::pandoc_to("latex")

table_format <- ifelse(is_docx, "huxtable", 'default')

# I don't know how Texas Instruments smart this is, but p-sure default DPI is 96.
# That's not a problem for LaTeX, but it looks not-so-great for Word.
# For Word, let's up that to 600. This should allow for cross-referencing 
# in {bookdown} while allowing for conditional DPI

if (is_latex) {
  conditional_dpi <- 96
  return(conditional_dpi)
} else {
  conditional_dpi <- 600
  return(conditional_dpi)
}

knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE, warning=FALSE,
                      fig.path='doc/figs/',
                      cache.path = 'doc/_cache/',
                      fig.width = 8.5, dpi = conditional_dpi,
                      fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      x2 = paste0(stringr::str_extract(x2, '.*/'), 
                                  "appendix-", 
                                  stringr::str_remove(x2, ".*/"))
                      if (file.rename(x, x2)) x2 else x
                      })

options(knitr.kable.NA = '')
```

\setcounter{table}{0}
\renewcommand*{\thetable}{A\arabic{table}}
```{r, results='asis'}
appendix_items <- readRDS("data/appendix_items.Rds")
EUI_final_DI_disc <- readRDS("data/EUI_final_DI_disc.Rds")

library(psychTools)
cat(fa2latex(appendix_items$`Factor analysis of economy`, heading = "Factor analysis of perception of economy", silent = TRUE, caption = "Factor analysis of perceptions of the economy."))
```

\setcounter{figure}{0}
\renewcommand*{\thefigure}{A\arabic{figure}}
```{r, fig.cap="Sensitivity to omitted variable bias for article-years (logged)", fig.height=8.5}
library(sensemakr)
appendix_items$"Fixed effect of non-interacted model" <- lm(opt_out_supp ~ article_disc_log + alternative_econ + gov_trust + left_right + left_right_sqr +
                          eurosceptic + gender + excl_identity + factor(country), weights = EUI_final_DI_disc$weight, 
                        data = EUI_final_DI_disc)
#Run sensemakr
appendix_items$"Sensitivity test of non-interacted model" <- sensemakr::sensemakr(appendix_items$`Fixed effect of non-interacted model`, 
                                treatment = "article_disc_log")
plot(appendix_items$"Sensitivity test of non-interacted model", type = "extreme")
```

```{r, fig.cap="Sensitivity to omitted variable bias for article-years (logged) X Euroscepticism", fig.height=8.5}
appendix_items$"Fixed effect of interacted model" <- lm(opt_out_supp ~ article_disc_log + alternative_econ + gov_trust + left_right + left_right_sqr +
                        eurosceptic + gender + excl_identity + article_disc_log*eurosceptic + factor(country), weights = EUI_final_DI_disc$weight, 
                      data = EUI_final_DI_disc)
appendix_items$"Sensitivity test of interacted model" <- sensemakr(appendix_items$`Fixed effect of interacted model`, 
                                  treatment = "article_disc_log:eurosceptic1")
plot(appendix_items$`Sensitivity test of interacted model`, type = "extreme")
```

```{r, fig.cap="Sensitivity to omitted variable bias for article-years (logged) X Euroscepticism in Scandinavia", fig.height=8.5}
nordic_data <- readRDS("data/nordic_data.Rds")

appendix_items$"Fixed effect of interacted model (Nordics)" <- lm(opt_out_supp ~ article_disc_log + alternative_econ + gov_trust + left_right + left_right_sqr +
                                                          eurosceptic + gender + excl_identity + article_disc_log*eurosceptic + factor(country), weights = nordic_data$weight, 
                                                        data = nordic_data)
appendix_items$"Sensitivity test of interacted model (Nordics)" <- sensemakr(appendix_items$`Fixed effect of interacted model (Nordics)`, 
                                                                   treatment = "article_disc_log:eurosceptic1")
plot(appendix_items$`Sensitivity test of interacted model (Nordics)`, type = "extreme")
```

```{r, results='asis'}
library(mice)
library(miceadds)
library(merTools)
library(broom)
library(broom.helpers)
main_impute <- readRDS("data/main_impute.Rds")

coef_map_regions2 <- c('(Intercept)' = 'Constant',
                          'article_disc_log:nordic_dummy1' = 'Article-years (logged) X Scandinavia',
                      'article_disc_log:CEE_dummy1' = 'Article-years (logged) X CEE',
                      "article_disc_log:eurosceptic1" = "Article-years (logged) X Eurosceptic",
                          'article_disc_log' = 'Article-years (logged)',
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                      'age' = "Age",
                      'income_scale' = "Income",
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eurosceptic1' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'excl_identity' = 'Exclusively national identity', 
                      "nordic_dummy" = "Scandinavia",
                      "CEE_dummy" = "Central and Eastern Europe")

modelsummary::modelsummary(main_impute$`Main model (Euroscepticism)`, fmt = 3, output = "latex", coef_map = coef_map_regions2, 
             stars = TRUE, title = "Imputed model of interaction between Euroscepticism and article-years (logged). Fixed country effects and SEs clustered at country level.")
```

```{r, results='asis'}
regional_impute <- readRDS("data/regional_impute.Rds")
coef_map_eurosceptic2 <- c('(Intercept)' = 'Constant',
                          'article_disc_log:eurosceptic1' = 'Article-years (logged) X Euroscepticism',
                          'article_disc_log' = 'Article-years (logged)',
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eurosceptic1' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'excl_identity' = 'Exclusively national identity')
modelsummary::modelsummary(regional_impute$`Article-years (logged) X regional dummies`, fmt = 3, output = "latex", coef_map = coef_map_regions2, 
             stars = TRUE, title = "Imputed models showing interactions of regional dummies and article-years (logged) by region. Fixed effects omitted due to collinearity. SEs clustered at country level.")
```

```{r, results='asis'}
region_mod <- readRDS("data/region_mod.Rds")
coef_map_eurosceptic2 <- c('(Intercept)' = 'Constant',
                          'article_disc_log:eurosceptic1' = 'Article-years (logged) X Euroscepticism',
                          'article_disc_log' = 'Article-years (logged)',
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                          'age' = "Age", 
                          'income_scale' = "Income",
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eurosceptic1' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'excl_identity' = 'Exclusively national identity')
modelsummary::modelsummary(region_mod, fmt = 3, output = "latex", coef_map = coef_map_regions2, 
             stars = TRUE, title = "Imputed models showing interactions of Euroscepticism and article-years (logged) by region") %>% kableExtra::kable_styling(latex_options = "scale_down")
```

```{r} 
coef_map_alteurosceptic <- c('(Intercept)' = 'Constant',
                          'article_disc_log:eu_democ_satisf' = 'Article-years (logged) X Satisfaction with EU democracy',
                          'article_disc_log' = 'Article-years (logged)',
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eu_democ_satisf' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'age' = "Age", 
                          'income_scale' = "Income",
                          'excl_identity' = 'Exclusively national identity')

modelsummary::modelsummary(appendix_items$`Alternative conceptualization of Euroscepticism`, fmt = 3, output = "latex", coef_map = coef_map_alteurosceptic, 
             stars = TRUE, title = "Model showing alternative conceptualization of Euroscepticism", vcov = ~country, gof_map = c("nobs", "adj.r.squared"))
```

```{r, fig.cap="Predicted values of support for constitutional differentiated integration as a function of an interaction between exposure to differentiated integration and satisfaction with EU democracy. Slopes for minimum, median and maximum values of satisfaction with EU democracy are shown. 95% CIs", fig.height=8.5}
library(sjPlot)
library(ggplot2)
pred_plot_interact_main <- plot_model(appendix_items$`Alternative conceptualization of Euroscepticism`, type = "pred", terms = c("article_disc_log", "eu_democ_satisf"), 
                                      colors = "bw", legend.title = "EU democracy satisfaction", vcov.fun = "vcovCL", vcov.args = list(cluster = EUI_final_DI_disc$country))

pred_plot_interact_main_final <-  pred_plot_interact_main + ggtitle("Support for constitutional differentiation", 
                                                                    subtitle = "Interaction between EU democracy satisfaction and differentiation") + labs(x = "Differentiation (article-years (logged))",
                                                                                                                                                           y = "Predicted value") + scale_linetype_manual(name = "EU democracy satisfaction",
                           values = c("dashed", "solid", "dotted"),
                           labels = c("Minimum", "Median", "Max")) + font_size(title = 15, labels.x = 15, labels.y = 15)
pred_plot_interact_main_final
```

```{r, fig.cap="Number of differentiations from EU primary law 1952-2020 (log-transformed and recency-weighted article-years) by EU member state.", fig.height=8.5}
bar_chart_articleyear <- ggplot(EUI_final_DI_disc, aes(x = factor(country), y = article_disc_log)) +
  geom_col(stat = "identity", position = "dodge")
bar_chart_articleyear + labs(x = "Country", y = "Discounted article-years (log-transformed)") +
  scale_x_discrete(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14),
                   labels = c("UK", "DK", "FI", "FR", "DE", "SE", "GR", "HU", "IT",
                              "LT", "NL", "PL", "RO", "ES")) +
  ggtitle(label = "Differentiated integration in the EU", 
          subtitle = "Primary law differentiations (article-years log-transformed and weighted)") + font_size(title = 15, labels.x = 15, labels.y = 15)
```

```{r, results='asis'}
appendix_items$"Southern Europe" <- lm(opt_out_supp ~ article_disc_log + market_europe + alternative_econ + gov_trust + left_right + left_right_sqr +
                           eurosceptic + gender + age + income_scale + excl_identity + south_dummy + south_dummy*article_disc_log, weights = EUI_final_DI_disc$weight, 
                         data = EUI_final_DI_disc)

coef_map_southern <- c('(Intercept)' = 'Constant',
                          'article_disc_log:south_dummy' = 'Article-years (logged) X Southern Europe',
                          'article_disc_log' = 'Article-years (logged)',
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eurosceptic1' = 'Eurosceptic',
                          'gender' = 'Gender',
                       'age' = "Age", 
                       'income_scale' = "Income",
                          'excl_identity' = 'Exclusively national identity', 
                      "south_dummy" = "Southern Europe")

modelsummary::modelsummary(appendix_items$`Southern Europe`, fmt = 3, output = "latex", coef_map = coef_map_southern, 
             stars = TRUE, title = "Model showing effect of exposure in Southern Europe (fixed effects omitted due to collinearity", vcov = sandwich::vcovCL(appendix_items$`Southern Europe`, ~country), gof_map = c("nobs", "adj.r.squared"))
```

```{r, fig.pos="H", fig.cap= "Support for differentiated integration in Southern Europe as an interaction between Euroscepticism and exposure to differentiated integration. 95% CIs.", fig.height=8.5}
pred_plot_southern <- plot_model(appendix_items$`Southern Europe`, type = "pred", terms = c("article_disc_log", "south_dummy"), 
                                 colors = "bw", legend.title = "Southern Europe", vcov.fun = "vcovCL", vcov.args = list(cluster = EUI_final_DI_disc$country))
pred_plot_southern_final <-  pred_plot_southern + ggtitle("Support for constitutional differentiation (Southern Europe)", subtitle = "Support for differentiation among Southern and non-Southern Europeans") + labs(x = "Differentiation (article-years (logged))",
                                                                                                                                 y = "Predicted value")
pred_plot_southern_final + font_size(title = 15, labels.x = 15, labels.y = 15) + scale_linetype_manual(name = "Southern Europe",
                           values = c("dashed", "solid"),
                           labels = c("Non-Southern", "Southern"))
```

```{r }
scandi_eurosceptic = list()
scandi_eurosceptic$"Scandinavia (three-way interactions)" <- lm(opt_out_supp ~ article_disc_log + market_europe + alternative_econ + gov_trust + left_right + left_right_sqr +
                             eurosceptic + gender + excl_identity + age + income_scale + nordic_dummy + article_disc_log*eurosceptic*nordic_dummy + factor(country), weights = EUI_final_DI_disc$weight, 
                           data = EUI_final_DI_disc)
```

```{r}

coef_map_nordic <- c('(Intercept)' = 'Constant',
                          'article_disc_log:eurosceptic1:nordic_dummy1' = 'Article-years (logged) X Eurosceptic X Scandinavia',
                          'article_disc_log' = 'Article-years (logged)',
                          'nordic_dummy1' = "Scandinavia",
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                     'age' = "Age", 
                     'income_scale' = "Income",
                          'left_right' = 'Left-right',
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eurosceptic1' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'excl_identity' = 'Exclusively national identity')

modelsummary::modelsummary(scandi_eurosceptic$"Scandinavia (three-way interactions)", fmt = 3, output = "latex", coef_map = coef_map_nordic,
             stars = TRUE, title = "Model showing three-way interaction between Scandinavian dummy, Eurosceptic and article-years (logged)", vcov = sandwich::vcovCL(scandi_eurosceptic$"Scandinavia (three-way interactions)", ~country), gof_map = c("nobs", "adj.r.squared"))
```

```{r}

main_mod_CSP = list()

main_mod_CSP$"Euroscepticism" <- lm(opt_out_supp ~ CSP_disc_log + market_europe + alternative_econ + gov_trust + left_right + left_right_sqr +
                                                         eurosceptic + gender + income_scale + age + excl_identity + CSP_disc_log*eurosceptic + factor(country), weights = EUI_final_DI_disc$weight, data = EUI_final_DI_disc)

coef_map_CSP <- c('(Intercept)' = 'Constant',
                          'CSP_disc_log:eurosceptic1' = 'Article-years (logged) X Euroscepticism',
                          "CSP_disc_log:nordic_dummy1" = "Article-years (logged) X Scandinavia", 
                          "CSP_disc_log:CEE_dummy1" = "Article-years (logged) X CEE",
                          "nordic_dummy1" = "Scandinavia",
                          "CEE_dummy1" = "Central and Eastern Europe",
                          'CSP_disc_log' = 'Article-years (logged)',
                          'market_europe' = 'Liberal economic values',
                          'alternative econ' = 'Perception of economy',
                          'gov_trust' = 'Support national government',
                          'left_right' = 'Left-right',
                          'left_right_sqr' = 'Left-right (sqr)',
                          'eurosceptic1' = 'Eurosceptic',
                          'gender' = 'Gender',
                          'age' = 'Age',
                          'income_scale' = 'Income',
                          'excl_identity1' = 'Exclusively national identity')

modelsummary::modelsummary(main_mod_CSP$"Euroscepticism", fmt = 3, output = "latex", coef_map = coef_map_CSP,
             stars = TRUE, title = "Model showing interaction between Euroscepticism and DI in the field of core state powers", vcov = sandwich::vcovCL(main_mod_CSP$"Euroscepticism", ~country), gof_map = c("nobs", "adj.r.squared"))
```

```{r, fig.pos="H", fig.cap="Predicted probability of support for constitutional differentiation as a function of the interaction between article-years (logged) in the field of core state powers and Euroscepticism. 95% country-clustered CIs with country fixed effects."}
library(sjPlot)
library(ggplot2)
pred_plot_eurosceptic_CSP <- plot_model(main_mod_CSP$`Euroscepticism`, type = "int", 
                                    colors = "bw", vcov.fun = "vcovCL", vcov.args = list(cluster = EUI_final_DI_disc$country))
pred_plot_eurosceptic_CSP_final <-  pred_plot_eurosceptic_CSP + ggtitle("Support for constitutional differentiation (exposure to CSP DI)", 
                                                                subtitle = "Support for differentiation among Eurosceptics and non-Eurosceptics") + labs(x = "Differentiation (article-years (logged))",
                                                                                                                     y = "Predicted value") + scale_linetype_manual(name = "Eurosceptic",
                           values = c("dashed", "solid", "dotted"),
                           labels = c("Non-Eurosceptic", "Eurosceptic"))
pred_plot_eurosceptic_CSP_final <- pred_plot_eurosceptic_CSP_final + font_size(title = 15, labels.x = 15, labels.y = 15)
pred_plot_eurosceptic_CSP_final
```